\chapter*{General Introduction}
\addcontentsline{toc}{chapter}{General Introduction}

Urban mobility has become a critical challenge in modern cities, particularly within corporate environments where large numbers of employees commute daily to a shared workplace. The increasing reliance on private vehicles has led to several issues, including traffic congestion, high transportation costs, and significant environmental impact due to excessive carbon emissions. In this context, carpooling has emerged as a promising solution to optimize vehicle usage and reduce the number of cars on the road.

However, despite its potential, traditional carpooling systems remain limited in their effectiveness. Most existing solutions rely on rigid matching mechanisms that require passengers to be located near a driver's fixed departure point. This constraint significantly reduces matching opportunities and does not reflect real-world commuting patterns, where employees are distributed across different geographic areas. Furthermore, these systems often lack intelligent coordination tools and do not exploit user behavior to improve the matching process.

In corporate environments, these limitations are even more pronounced, as commuting patterns are often repetitive but geographically dispersed. Employees may share similar destinations but differ in departure locations and schedules, making rigid matching approaches inefficient. As a result, there is a need for more flexible and intelligent systems capable of adapting to real-world constraints while maintaining usability and efficiency.

This project was carried out within the framework of an end-of-studies internship in a corporate environment, where the need for an efficient and intelligent commuting solution was clearly identified. The objective was not only to design a theoretical model but also to develop a functional and scalable platform that can be deployed in a real organizational context.

To address the identified limitations, this work proposes the design and implementation of an intelligent carpooling platform tailored for enterprise use. The system introduces a flexible and user-centric approach that combines geospatial analysis, real-time communication, and machine learning techniques to enhance ride-sharing efficiency and user experience.

By integrating these components, the platform aims to transform traditional carpooling into a dynamic and adaptive system, capable of responding to user needs in real time. This shift from static to intelligent coordination represents a key step toward more efficient and sustainable mobility solutions within organizations.

A key innovation of this platform lies in its smart matching system, which is based on route overlap analysis rather than fixed origin proximity. By analyzing a driver's route and identifying potential pickup points along it, the system enables dynamic and flexible passenger pickup, thereby significantly increasing the likelihood of successful matches. This approach is implemented using deterministic geospatial algorithms, ensuring efficiency, scalability, and interpretability.

In addition to this rule-based intelligence, the platform integrates a machine learning subsystem designed to analyze historical commuting data and detect recurring travel patterns. By leveraging classification models, the system predicts the probability of a user's commute at a given time and provides proactive suggestions, allowing users to interact with the platform more effectively.

The combination of these two approaches — deterministic geospatial matching and probabilistic prediction — allows the system to balance precision and adaptability. This hybrid design enhances both system performance and user engagement, making the platform more responsive to real-world usage scenarios.

The proposed solution is implemented as a full-stack application, combining a Flask-based backend, a Flutter mobile frontend, and a PostgreSQL database. It also integrates external geospatial services such as OpenStreetMap and OSRM for route computation and location processing. Real-time features are supported through WebSocket communication, enabling live tracking and dynamic system updates.

The main objective of this work is to develop a robust, intelligent, and scalable carpooling platform that overcomes the limitations of traditional solutions. More specifically, the project aims to:
\begin{itemize}
    \item Improve ride-matching flexibility through route-based dynamic pickup.
    \item Enhance user experience via real-time communication and notifications.
    \item Introduce predictive capabilities using machine learning techniques.
    \item Provide a complete and deployable full-stack system.
\end{itemize}

To achieve these objectives, an Agile development methodology was adopted, enabling iterative design, implementation, and validation of system components throughout the internship period.

The development process required careful consideration of multiple constraints: scalability, real-time performance, usability, and integration with external services. Particular attention was given to maintainability and extensibility, enabling future improvements such as traffic data integration and more sophisticated learning models.

This work bridges the gap between theoretical models and real-world deployment, delivering a complete operational solution. By combining rule-based algorithms with machine learning techniques, the system achieves both interpretability and adaptability to user behavior.

This report is structured as follows:

\begin{itemize}

    \item \textbf{Chapter 1: Project Initiation} \\
    This chapter introduces the enterprise context and the commuting problem, analyzes existing solutions and their limitations, and presents the proposed solution. It also outlines the adopted Agile development methodology based on the Scrum framework, providing an overview of the project organization and development approach.

    \item \textbf{Chapter 2: System Architecture and Design} \\
    This chapter presents the overall system architecture of the platform. It describes the layered structure, the main system components, their interactions, and the key architectural design decisions. The system is also modeled through global use case and class diagrams to provide a clear understanding of its structure before implementation.

    \item \textbf{Chapter 3: User Management and Authentication} \\
    This chapter focuses on the implementation of user management functionalities, including enterprise authentication, domain validation, and profile management. It also introduces administrative features such as user monitoring and account control, ensuring secure and controlled access to the system.

    \item \textbf{Chapter 4: Ride Management and Real-Time System} \\
    This chapter presents the implementation of the core ride management functionalities, including ride creation, search, booking, and lifecycle management. It also covers the integration of real-time communication using WebSocket technology for live updates and user interaction.

    \item \textbf{Chapter 5: Intelligent System (Matching and Machine Learning)} \\
    This chapter describes the intelligent components of the system, including the route-based matching algorithm and the machine learning model for commute prediction. It explains how these components enhance the platform by providing optimized ride suggestions and proactive recommendations.

    \item \textbf{Chapter 6: Testing, Validation and Results} \\
    This chapter presents the validation of the system through testing and performance evaluation. It highlights the reliability of the platform, assesses the effectiveness of implemented functionalities, and demonstrates compliance with the defined requirements.

\end{itemize}

This work contributes to the field of intelligent transportation systems by proposing a hybrid approach that combines deterministic algorithms and machine learning techniques to improve carpooling efficiency and usability in real-world enterprise environments.