\chapter{System Evaluation and Validation}

\section{Introduction}

This chapter presents the final stage of the project, dedicated to the global evaluation and validation of the proposed platform. Unlike previous sprints, where validation focused on individual functionalities, this phase adopts a broader perspective by assessing the behavior of the system as a fully integrated solution.

The objective is to verify that the different components — ride management, real-time communication, intelligent matching, and predictive mechanisms — operate coherently under realistic usage conditions. Particular attention is given to interaction consistency, synchronization reliability, and the practical contribution of intelligent features to the overall user experience.

% =========================================================

\section{Sprint 6: Final Validation and System Evaluation}

\subsection{Sprint Objective}

The objective of this sprint is to validate the platform as a complete and integrated environment. This includes verifying end-to-end workflows, evaluating synchronization across multiple users, assessing the effectiveness of intelligent mechanisms, and analyzing system robustness under both normal and edge-case conditions.

% ------------------------

\subsection{Sprint Backlog}

The backlog of this sprint focuses on system-wide validation rather than isolated feature verification. The evaluation phase combines integration testing, synchronization analysis, intelligent feature assessment, and robustness verification.

\begin{table}[H]
\centering
\renewcommand{\arraystretch}{1.2}
\setlength{\tabcolsep}{5pt}

\begin{tabular}{|p{0.75\textwidth}|c|}
\hline
\textbf{User Story} & \textbf{Priority} \\
\hline
As a system, I want to validate complete end-to-end workflows across all modules to ensure global system coherence & High \\
\hline
As a system, I want to verify integration between ride management, real-time communication, and intelligent components to ensure seamless interaction & High \\
\hline
As a system, I want to evaluate real-time synchronization across multiple clients to ensure consistent shared state & High \\
\hline
As a system, I want to assess the effectiveness of the smart matching mechanism in realistic ride scenarios & High \\
\hline
As a system, I want to evaluate the relevance and usefulness of machine learning predictions in practical contexts & High \\
\hline
As a system, I want to analyze system responsiveness and perceived latency during typical user interactions & Medium \\
\hline
As a system, I want to test robustness against edge cases and invalid operations to ensure stability & High \\
\hline
As a system, I want to leverage observability tools (logs, events, dashboard) to verify system behavior and trace execution & Medium \\
\hline
As a system, I want to identify limitations and areas for improvement to guide future enhancements & Medium \\
\hline
\end{tabular}

\caption{Sprint 6 Backlog — System Evaluation}
\end{table}

The sprint therefore serves as a final verification phase intended to confirm that the platform behaves consistently when all subsystems interact simultaneously.

% =========================================================

\section{Evaluation Methodology}

The evaluation follows a scenario-based methodology designed to reproduce realistic system usage rather than isolated functional tests. Each scenario involves several components interacting simultaneously, enabling observation of platform behavior under practical operating conditions.

The evaluation primarily focuses on:

\begin{itemize}
    \item end-to-end ride workflow execution,
    \item multi-user synchronization and interaction,
    \item real-time propagation of events,
    \item intelligent decision-making behavior.
\end{itemize}

The purpose of this methodology is not limited to verifying functional correctness. It also evaluates interaction consistency, synchronization reliability, responsiveness, and the practical usefulness of intelligent system features.

This broader perspective is important because the proposed platform relies heavily on interactions between distributed and asynchronous components, where isolated testing alone would not sufficiently reflect real operational behavior.

% =========================================================

\section{End-to-End System Validation}

A complete ride execution scenario is performed, covering the full workflow from ride creation to ride completion. The sequence includes:

\begin{itemize}
    \item creation of a ride by the driver with predefined parameters,
    \item discovery and reservation request by a passenger,
    \item validation and acceptance of the request,
    \item progression of the ride through its lifecycle states.
\end{itemize}

Throughout the process, the platform successfully enforces all validation constraints and business rules. Seat availability remains consistent, invalid operations are rejected before persistence, and lifecycle transitions occur in a controlled manner.

The evaluation also confirms that synchronization between mobile clients and backend services remains coherent during ride execution. State updates are propagated correctly across connected users, ensuring that all participants observe a consistent system state in real time.

These observations demonstrate that the backend-centered architecture effectively preserves system integrity across complex multi-stage workflows while maintaining reliable interaction between distributed components.
---

\section{Real-Time System Evaluation}

The real-time communication layer is evaluated through simultaneous interaction between multiple connected clients operating on the same ride session.

Whenever ride states change or driver location updates are emitted, the corresponding events are propagated to all subscribed users with an observed latency below 0.5~s. Compared to traditional polling-based synchronization, the event-driven approach reduces perceived update latency by more than 75\%, while also eliminating unnecessary repetitive requests.

The evaluation further confirms that connected clients maintain a coherent shared state throughout ride execution. Updates are reflected immediately across interfaces without requiring manual refresh operations, ensuring continuous synchronization between participants.

Connection interruptions are also handled correctly. When a client reconnects after temporary disconnection, the system restores synchronization using the latest backend state, allowing interaction continuity without observable inconsistencies.

These observations validate that the WebSocket-based architecture provides efficient and reliable real-time coordination under realistic multi-user conditions.

% =========================================================

\section{Intelligent System Evaluation}

\subsection{Smart Matching Evaluation}

The smart matching mechanism is evaluated through scenarios in which passengers are located along the driver's route rather than exactly at the origin or destination points.

The evaluation confirms that the system successfully identifies feasible ride opportunities by projecting passenger locations onto driver trajectories and computing acceptable detour costs. Unlike strict origin--destination matching approaches, the proposed trajectory-based strategy supports intermediate insertion points while preserving route continuity.

Experimental observations indicate that the route-based approach increases matching opportunities by approximately 27 percentage points compared to rigid endpoint matching, while maintaining geographically coherent routes.

These results validate the effectiveness of the corridor-based matching strategy in handling realistic spatial constraints and improving ride discovery flexibility.

% ------------------------

\subsection{Machine Learning Evaluation}

The machine learning component is evaluated with respect to both prediction relevance and classification performance.

The evaluation confirms that the predictive layer successfully identifies recurring commuting behaviors and generates ride suggestions that remain aligned with observed user habits. The Random Forest classifier achieves an F1-score of 0.82, with precision and recall values of 0.79 and 0.85 respectively, indicating balanced predictive performance.

The results also show that the model remains capable of capturing structured behavioral patterns despite relying on synthetic training data. Generated predictions remain temporally and spatially coherent, supporting the feasibility of predictive assistance within the proposed platform.

Although the evaluation is conducted in a controlled environment, the observed behavior demonstrates that the predictive layer can effectively complement the rule-based ride management workflow.

% =========================================================

\section{Experimental Evaluation}

\subsection{Evaluation Metrics}

The following quantitative metrics are used to assess system performance:

\begin{itemize}
    \item \textbf{Matching Success Rate} -- proportion of passenger requests for which at least one feasible driver match is identified.
    
    \item \textbf{Average Pickup Deviation} -- mean geographic distance (km) between the passenger's declared origin and the computed pickup point.
    
    \item \textbf{Real-Time Latency} -- elapsed time (seconds) between a backend state-changing event and its reception by connected clients.
    
    \item \textbf{Prediction Accuracy} -- overall correctness of the commute prediction model.
    
    \item \textbf{Precision / Recall / F1-score} -- standard evaluation measures for the binary commute-prediction task.
\end{itemize}

These metrics provide complementary perspectives on system behavior, covering synchronization efficiency, spatial matching quality, and predictive performance.

% ------------------------

\subsection{Baseline System}

To isolate the contribution of the intelligent mechanisms, a baseline system is defined for comparison purposes.

The baseline implements:

\begin{itemize}
    \item strict origin--destination matching, where passenger and driver origins must coincide,
    
    \item no route-based flexibility or spatial corridor analysis,
    
    \item no predictive layer; ride discovery relies entirely on manual user interaction.
\end{itemize}

This configuration reflects the behavior of conventional ride-sharing systems and serves as a reference point for evaluating the impact of the proposed architecture.

% ------------------------

\subsection{Experimental Setup}

The evaluation is conducted using a synthetic dataset representing enterprise commuting behavior. The dataset includes multiple users, predefined home-to-work trajectories, and departure-time distributions concentrated around typical commuting periods.

Both the baseline and the proposed platform are evaluated under identical experimental conditions. The same user profiles, route structures, and temporal distributions are supplied to both systems to ensure fair comparison.

This controlled setup guarantees that observed performance differences result from architectural improvements rather than variations in input data.

% ------------------------

\subsection{Results}

Table~\ref{tab:experimental_results} summarizes the comparative evaluation results.

\begin{table}[H]
\centering
\renewcommand{\arraystretch}{1.3}

\begin{tabular}{|l|c|c|}
\hline
\textbf{Metric} & \textbf{Baseline} & \textbf{Proposed System} \\
\hline
Matching Success Rate & 45\% & 72\% \\
\hline
Average Pickup Deviation (km) & 0.0 & 1.8 \\
\hline
Real-Time Latency (s) & $\sim$2.1 & $<$0.5 \\
\hline
Prediction Accuracy & -- & 0.82 \\
\hline
Precision & -- & 0.79 \\
\hline
Recall & -- & 0.85 \\
\hline
F1-score & -- & 0.82 \\
\hline
\end{tabular}

\caption{Experimental Results -- Baseline vs. Proposed System}
\label{tab:experimental_results}

\end{table}
\subsection{Results Interpretation}

The experimental results indicate that the proposed platform performs significantly better than the baseline across all evaluated dimensions.

The matching success rate increases from 45\% to 72\%, representing a substantial improvement in ride discovery capability. This gain is primarily explained by the trajectory-based matching strategy, which evaluates spatial overlap along the complete route rather than relying on strict origin--destination coincidence.

At the same time, the average pickup deviation remains limited to 1.8~km, indicating that the additional flexibility introduced by corridor-based matching does not generate excessive detours for drivers. The optimization mechanism therefore succeeds in balancing matching expansion with route coherence.

The evaluation also highlights a major improvement in synchronization responsiveness. Real-time event latency decreases from approximately 2.1~s to below 0.5~s, confirming the efficiency of the WebSocket-based communication layer compared to polling-oriented approaches. This reduction contributes directly to smoother interaction and improved user perception of system responsiveness.

Regarding predictive behavior, the machine learning component achieves an F1-score of 0.82, with precision and recall values of 0.79 and 0.85 respectively. These results suggest that the classifier effectively captures recurring commuting habits while maintaining an acceptable balance between missed opportunities and irrelevant suggestions.

Overall, the evaluation confirms that combining route-based matching, real-time synchronization, and predictive assistance leads to measurable improvements in both operational efficiency and interaction quality.

The results also reveal an important trade-off between flexibility and route deviation. Increasing matching tolerance improves the probability of identifying feasible rides, but introduces moderate pickup displacement as passengers are inserted along trajectories rather than at exact endpoints.

However, the observed deviation remains compatible with typical urban commuting distances, confirming that the proposed strategy remains practical in realistic deployment conditions.

% =========================================================

\subsection{Key Findings}

The experimental evaluation leads to the following observations:

\begin{itemize}
    \item \textbf{Matching effectiveness}: the proposed platform identifies substantially more feasible ride opportunities than the baseline, increasing the success rate from 45\% to 72\%.
    
    \item \textbf{Route preservation}: the average pickup deviation remains limited to 1.8~km, indicating that increased flexibility does not significantly disrupt driver routes.
    
    \item \textbf{Real-time responsiveness}: event propagation latency remains below 0.5~s, representing a reduction greater than 75\% compared to polling-based synchronization.
    
    \item \textbf{Prediction reliability}: the Random Forest model achieves an F1-score of 0.82, demonstrating stable predictive capability for recurring commuting behavior.
    
    \item \textbf{System consistency}: invalid operations are systematically rejected at the backend level, and concurrent interactions do not produce observable state inconsistencies.
    
    \item \textbf{Observability support}: logs, events, and monitoring indicators provide continuous visibility into ride execution and intelligent decision-making processes.
\end{itemize}

% =========================================================

\section{System Robustness and Observability}

The platform is evaluated against multiple edge-case scenarios, including invalid booking attempts, concurrent requests, and inconsistent user actions.

The evaluation confirms that invalid operations are consistently rejected by backend validation mechanisms before affecting persistent system state. Concurrent interactions also remain correctly synchronized, particularly with respect to seat allocation and lifecycle management.

The observability layer further contributes to validation by providing detailed visibility into internal system activity. Logs, system events, and monitoring indicators enable precise tracking of ride execution, synchronization events, and matching decisions.

This traceability simplifies debugging and helps verify that distributed interactions remain coherent throughout execution.

% =========================================================

\section{Discussion and Limitations}

Although the evaluation demonstrates strong system coherence and performance, several limitations remain.

First, the machine learning component relies on synthetic data, which cannot fully reproduce the variability of real-world commuting behavior. Integrating real usage data in future deployments would improve predictive realism and allow continuous adaptation to evolving user habits.

Second, the evaluation primarily focuses on functional correctness and interaction behavior rather than large-scale stress testing. Scalability under high concurrency therefore remains an open consideration.

A preliminary analysis suggests that a significant increase in concurrent users would mainly impact two components. The first concerns the matching pipeline, where corridor-based filtering complexity grows with the number of active rides and pending requests. Without advanced spatial indexing techniques, processing latency could increase beyond the current near real-time behavior.

The second concerns the real-time communication layer. Since WebSocket sessions currently maintain server-side connection state, large-scale deployment would require distributed synchronization mechanisms and shared event propagation infrastructure to preserve consistency across multiple backend instances.

These observations indicate that the current architecture is well adapted to enterprise-scale deployment involving tens or low hundreds of simultaneous users, while larger deployments would require additional horizontal scaling strategies.

Finally, although the synchronization layer performs efficiently under controlled conditions, network instability may still affect responsiveness in environments with unreliable connectivity.

% =========================================================

\section{Conclusion}

This chapter demonstrated through quantitative and behavioral evaluation that the proposed platform significantly improves upon conventional ride-sharing approaches.

The experimental results confirm three principal outcomes:

\begin{itemize}
    \item route-based matching substantially increases the number of feasible ride opportunities,
    
    \item WebSocket-driven synchronization considerably reduces interaction latency,
    
    \item the predictive layer successfully anticipates recurring commuting behavior through machine learning techniques.
\end{itemize}

Beyond these isolated metrics, the evaluation also confirms that the backend-centered architecture maintains consistent system behavior under concurrent interactions, invalid operations, and edge-case scenarios.

The observability mechanisms provide complete traceability of ride lifecycle transitions and intelligent decision processes, supporting both runtime validation and post-execution analysis.

Overall, the evaluation validates the effectiveness of combining real-time communication, intelligent spatial matching, and predictive assistance within a unified enterprise ride-sharing platform.